---
title: "Live Session 3"
author: "Jeremy Otsap"
date: "May 21, 2019"
output: 
  html_document:
    keep_md: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 2

A) Loading titanic.csv into data frame

```{r, echo = T}
read.csv("https://raw.githubusercontent.com/jotsap/DDS6306_Unit03/master/titanic/titanic.csv", sep = ",", header = T) -> titanic.df
```

B) Output the respective count of females and males aboard the Titanic

```{r, echo = T}
#all 'female' rows
titanic.df[titanic.df$Sex == 'female',] -> titanic.female
print("female passengers")
nrow(titanic.female) 

#all 'male' rows
titanic.df[titanic.df$Sex == 'male',] -> titanic.male
print("male passengers")
nrow(titanic.male) 
```


### Pie Chart of Male vs Female
```{r, echo = T}
#Pie Chart male vs femaile
titanic.table <- table(titanic.df$Sex)
titanic.labels <- paste(names(titanic.table), "\n", titanic.table, sep="")
pie(titanic.table, labels = titanic.labels, main = "Pie Chart of Gender") 



```

C) Means of Age, Fare, and Survival

```{r, echo = T}
# ignoring NA values
sapply(titanic.df[,c(6,10,2)], mean, na.rm=T) -> titanic.mean
titanic.mean
```


## Question 3

A) Creating the Helper function
```{r, echo = T}
#loading sleep data set
sleep.df <- read.csv("http://talklab.psy.gla.ac.uk/L1_labs/lab_1/homework/sleep_data_01.csv", sep = ",", header = T)

#function for 
#median of Age
#range of Duration
#mean of RSES
#std dev of RSES

sleephelp <- function(a,b,c,d) {
  aa <- median(a, na.rm = T)
  bb <- range(b, na.rm = T)
  cc <- mean(c, na.rm = T)
  dd <- sd(d, na.rm = T)
  output <- c(
    Age_Median = aa,
    Duration_Range = bb,
    RSES_Mean = cc,
    RSES_StdDev = dd
  )
  return( output)
}

sleephelp(sleep.df$Age,sleep.df$Duration,sleep.df$RSES,sleep.df$RSES)


```


B, C, E) Report Output
```{r, echo = T}

#report function
#median of age
#RSES Mean / 5
#RSES Std Dev / 5
#Duration Max - Min

sleepreport <- function(a,b,c) {
  aa <- median(a, na.rm = T)
  bb <- mean(b, na.rm = T) / 5
  cc <- sd(b, na.rm = T) / 5
  dd <- max(c, na.rm = T) - min(c, na.rm = T)
  output <- c(
    MedianAge = aa,
    SelfEsteem = bb,
    SE_SD = cc,
    DurationRange = dd
  )
  outputframe <- data.frame(output)
  return(outputframe)
}
sleepreport(sleep.df$Age,sleep.df$RSES,sleep.df$Duration)

```




## Question 4
A) Loading fivethirtyeight package & listing all data sets within package
```{r, echo = T}

#install fivethirtyeight packages
library(fivethirtyeight)

# To see a list of all data sets:
data(package = "fivethirtyeight")
```

B) Loading 22nd data set "college_recent_grads" into data frame

```{r, echo = T}
#loading 22nd data set "college_recent_grads" from fivethirtyeight package
df <- college_recent_grads
head(df)
```

C) Detailed view of data sets w/in fivethirtyeight package
URL is http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/
```{r, echo = T}

# To see a more detailed list of all data sets, see the package vignette:
vignette("fivethirtyeight", package = "fivethirtyeight") 

# URL is http://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/

```

D) Dimensions of dataframe

```{r, echo = T}
#dimensions of df
dim(df)

#column names of df
colnames(df)

#alternatively
str(df)

```



## Question 5

A) Column names and number of colums
```{r, echo = T}
# column names
colnames(df)

# number of columns
length(colnames(df))

```


B) Unique factors for major_category column in df
```{r, echo = T}

#much simpler to use count() from dplyr
library(dplyr)
count(df, major_category) -> major_count
major_count


```


C) Barplot of Graduates by Major

```{r, echo = T}

# uses rainbow() function to generate as many colors as major categories
# then legend will associate those colors to the respective major_category

barplot(
  major_count$n,
  col = rainbow(16), 
  main = "Graduates by Major",
  horiz = T,
  legend.text = major_count$major_category,
  xlab = "Num of Graduates",
  ylab = "Graduate Major"
  )

```


D) Export to CSV without row labels


```{r, echo = T}

write.csv(
  df, 
  file = "fivethirtyeight/fivethirtyeight.csv", 
  append = F, 
  sep = ",", 
  row.names = F
  )

```



## Question 6

Github repo located here:
https://github.com/jotsap/DDS6306_Unit03 

